Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-11-07 Robustness-Methods-and-Results.Rmd
ref fix

As with the main analyses, full surveys are available in our OSF repository \citep{KreienkampMasked2022a} and the full data description is available in Online Supplementary Material A. Correlations and descriptive statistics of the included variables are available in Table \ref{tab:descrFullWide} and Table \ref{tab:descrOutWide}.

@@

To build further confidence in our results, we assessed two additional models that might offer alternative explanations. First, to ensure that the effect of core need fulfillment is specific to an actual contact, we compared the effect to core need fulfillment in situations without an intergroup contact. For this, we analyzed the generalized situational core need fulfillment (either during a contact or about the daytime in general) and tested whether the effect differed during experience sampling measurements with and without outgroup contacts. We found no main effect of core need fulfillment (random slopes model; `r MlCoeffLatex(lmeMdl = mdlWorker$lmeSlopesAttCoreInt, lmerCI = mdlWorker$lmerSlopesAttCoreIntCI, varName = "keyMotiveFulfilled_cwc")`) but a significant interaction effect of core need fulfillment and outgroup contact (`r MlCoeffLatex(lmeMdl = mdlWorker$lmeSlopesAttCoreInt, lmerCI = mdlWorker$lmerSlopesAttCoreIntCI, varName = "keyMotiveFulfilled_cwc:OutgroupInteraction")`; also see Table \ref{tab:robustnessTblLong} and Figure \ref{fig:Robustness}). Together with a significant main effect of having an outgroup contact, this indicates that it is not key need fulfillment in general --- but only key need fulfillment during an outgroup contact that predicts more positive outgroup attitudes.

@@

In a final step, we controlled for other fundamental psychological needs during the contact. We focus on the three commonly considered self-determination needs (SDT): competence, autonomy, and relatedness. We find that the core need fulfillment adds significantly above a model with only the self-determination theory needs (random slopes models; `r workerSdtCompChiSq`). We also find that next to relatedness (`r MlCoeffLatex(lmeMdl = mdlWorkerOut$Att$lmeSlopesCoreSdt, lmerCI = mdlWorkerOut$Att$lmerSlopesCoreSdtCI, varName = "relatedness_1_cwc")`), the core need explains the most variance in outgroup attitudes after an outgroup contact (`r MlCoeffLatex(lmeMdl = mdlWorkerOut$Att$lmeSlopesCoreSdt, lmerCI = mdlWorkerOut$Att$lmerSlopesCoreSdtCI, varName = "keymotive_fulfillemt_1_cwc")`). When compared to the model with only the SDT needs, the core need fulfillment flexibly takes on some of the explained variance of all three fundamental needs (competence and autonomy needs turning non-significant; all \textit{b} < `r coef(summary(mdlWorkerOut$Att$lmeSlopesCoreSdt))[c("competence_1_cwc", "autonomy_1_cwc"),"Value"] %>% max %>% round(2) %>% format(nsmall=2)`, all \textit{p} > `r coef(summary(mdlWorkerOut$Att$lmeSlopesCoreSdt))[c("competence_1_cwc", "autonomy_1_cwc"),"p-value"] %>% min %>% round(3) %>% format(nsmall=3)`). For full results see see Table \ref{tab:robustnessTblLong}, Figure \ref{fig:Robustness}, and Online Supplementary Material A. There is, thus, considerable evidence lending confidence to the stability and relevance of psychological need fulfillment as a predictor of positive outgroup attitudes for natural intergroup contacts.

@@

Also for Study 2, we checked for alternative models. First, when considering generalized situational core need fulfillment together with whether an intergroup contact took place, we find that there is only a minuscule main effect of core need fulfillment (random slopes model; `r MlCoeffLatex(lmeMdl = mdlStudent$lmeSlopesAttCoreInt, lmerCI = mdlStudent$lmerSlopesAttCoreIntCI, varName = "KeyNeedFullfillment_cwc")`) but a stronger interaction effect of core need fulfillment and outgroup contact (`r MlCoeffLatex(lmeMdl = mdlStudent$lmeSlopesAttCoreInt, lmerCI = mdlStudent$lmerSlopesAttCoreIntCI, varName = "KeyNeedFullfillment_cwc:OutgroupInteraction")`). Together with a significant main effect of having an outgroup contact (`r MlCoeffLatex(lmeMdl = mdlStudent$lmeSlopesAttCoreInt, lmerCI = mdlStudent$lmerSlopesAttCoreIntCI, varName = "OutgroupInteraction")`), this indicates that it is not key need fulfillment in general --- but key need fulfillment during an outgroup contact that predicts more positive outgroup attitudes. This finding is consistent with the results of the previous study, albeit with a slightly weaker effect (likely because of the large number of measurements that did not include an outgroup interaction; For full results see Table \ref{tab:intergroupNeedsTblLong} and for visual comparison see Figure \ref{fig:AllportNeedFulfillment}).

@@

In a final step, we again checked whether during the interaction the core situational need remains a meaningful predictor even when taking other fundamental psychological needs into account. We find that the core need fulfillment adds significantly above a model with only the self-determination theory needs (random slopes models; `r studentSdtCompChiSq`). We find that the core need fulfillment explained the most variance in outgroup attitudes after an outgroup contact (`r MlCoeffLatex(lmeMdl = mdlStudentOut$Att$lmeSlopesCoreSdt, lmerCI = mdlStudentOut$Att$lmerSlopesCoreSdtCI, varName = "KeyNeedFullfillment_cwc")`). When compared to the model with only the SDT needs, the core need fulfillment flexibly takes on some of the explained variance of all of the three fundamental needs. However, different from the first study, relatedness (`r MlCoeffLatex(lmeMdl = mdlStudentOut$Att$lmeSlopesCoreSdt, lmerCI = mdlStudentOut$Att$lmerSlopesCoreSdtCI, varName = "Relatedness_cwc")`) and autonomy (`r MlCoeffLatex(lmeMdl = mdlStudentOut$Att$lmeSlopesCoreSdt, lmerCI = mdlStudentOut$Att$lmerSlopesCoreSdtCI, varName = "Autonomy_cwc")`) also predicted positive outgroup attitudes in this larger sample. For full results see Table \ref{tab:intergroupNeedsTblLong}, Figure \ref{fig:Robustness}, and Online Supplementary Material A. This means that also within this second sample, the fulfillment of psychological needs during intergroup contact remained a key predictor of positive outgroup attitudes, even when taking into account several alternative models.

@@

In a final step, we again checked whether during the interaction the core situational need remains a meaningful predictor even when taking other fundamental psychological needs into account. We find that the core need fulfillment adds additional variance above a model with only the self-determination theory needs (random slopes models; `r medicalSdtCompChiSq`). We find that the core need explains the most variance in outgroup attitudes after an outgroup contact (`r MlCoeffLatex(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreSdt, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreSdtCI, varName = "KeyNeedFulfillment_cwc")`). When compared to the model with only the SDT needs, the core need fulfillment flexibly takes on some of the explained variance of all of the three fundamental needs. However, similar to the previous study, in this large sample relatedness (`r MlCoeffLatex(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreSdt, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreSdtCI, varName = "Relatedness_cwc")`), competence (`r MlCoeffLatex(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreSdt, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreSdtCI, varName = "Competence_cwc")`) and autonomy (`r MlCoeffLatex(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreSdt, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreSdtCI, varName = "Autonomy_cwc")`) each also predicted positive outgroup attitudes independently. This being said, the regression coefficient for the core need was three times larger (with all scaling being equal). For full results see Table \ref{tab:robustnessTblLong} and Figure \ref{fig:Robustness} as well as Online Supplementary Material A. Across all three studies, psychological need fulfillment, thus, remained a robust and flexible predictor of positive outgroup attitudes.

>>>>

As with the main analyses, full surveys are available in our OSF repository \citep{KreienkampMasked2022a} and the full data description is available in Online Supplementary Material A. Correlations and descriptive statistics of the included variables are available in \tblref{tab:descrFullWide} and \tblref{tab:descrOutWide}.

@@

To build further confidence in our results, we assessed two additional models that might offer alternative explanations. First, to ensure that the effect of core need fulfillment is specific to an actual contact, we compared the effect to core need fulfillment in situations without an intergroup contact. For this, we analyzed the generalized situational core need fulfillment (either during a contact or about the daytime in general) and tested whether the effect differed during experience sampling measurements with and without outgroup contacts. We found no main effect of core need fulfillment (random slopes model; `r MlCoeffLatex(lmeMdl = mdlWorker$lmeSlopesAttCoreInt, lmerCI = mdlWorker$lmerSlopesAttCoreIntCI, varName = "keyMotiveFulfilled_cwc")`) but a significant interaction effect of core need fulfillment and outgroup contact (`r MlCoeffLatex(lmeMdl = mdlWorker$lmeSlopesAttCoreInt, lmerCI = mdlWorker$lmerSlopesAttCoreIntCI, varName = "keyMotiveFulfilled_cwc:OutgroupInteraction")`; also see \tblref{tab:robustnessTblLong} and \fgrref{fig:Robustness}). Together with a significant main effect of having an outgroup contact, this indicates that it is not key need fulfillment in general --- but only key need fulfillment during an outgroup contact that predicts more positive outgroup attitudes.

@@

In a final step, we controlled for other fundamental psychological needs during the contact. We focus on the three commonly considered self-determination needs (SDT): competence, autonomy, and relatedness. We find that the core need fulfillment adds significantly above a model with only the self-determination theory needs (random slopes models; `r workerSdtCompChiSq`). We also find that next to relatedness (`r MlCoeffLatex(lmeMdl = mdlWorkerOut$Att$lmeSlopesCoreSdt, lmerCI = mdlWorkerOut$Att$lmerSlopesCoreSdtCI, varName = "relatedness_1_cwc")`), the core need explains the most variance in outgroup attitudes after an outgroup contact (`r MlCoeffLatex(lmeMdl = mdlWorkerOut$Att$lmeSlopesCoreSdt, lmerCI = mdlWorkerOut$Att$lmerSlopesCoreSdtCI, varName = "keymotive_fulfillemt_1_cwc")`). When compared to the model with only the SDT needs, the core need fulfillment flexibly takes on some of the explained variance of all three fundamental needs (competence and autonomy needs turning non-significant; all \textit{b} < `r coef(summary(mdlWorkerOut$Att$lmeSlopesCoreSdt))[c("competence_1_cwc", "autonomy_1_cwc"),"Value"] %>% max %>% round(2) %>% format(nsmall=2)`, all \textit{p} > `r coef(summary(mdlWorkerOut$Att$lmeSlopesCoreSdt))[c("competence_1_cwc", "autonomy_1_cwc"),"p-value"] %>% min %>% round(3) %>% format(nsmall=3)`). For full results see see \tblref{tab:robustnessTblLong} and \fgrref{fig:Robustness}, and Online Supplementary Material A. There is, thus, considerable evidence lending confidence to the stability and relevance of psychological need fulfillment as a predictor of positive outgroup attitudes for natural intergroup contacts.

@@

Also for Study 2, we checked for alternative models. First, when considering generalized situational core need fulfillment together with whether an intergroup contact took place, we find that there is only a minuscule main effect of core need fulfillment (random slopes model; `r MlCoeffLatex(lmeMdl = mdlStudent$lmeSlopesAttCoreInt, lmerCI = mdlStudent$lmerSlopesAttCoreIntCI, varName = "KeyNeedFullfillment_cwc")`) but a stronger interaction effect of core need fulfillment and outgroup contact (`r MlCoeffLatex(lmeMdl = mdlStudent$lmeSlopesAttCoreInt, lmerCI = mdlStudent$lmerSlopesAttCoreIntCI, varName = "KeyNeedFullfillment_cwc:OutgroupInteraction")`). Together with a significant main effect of having an outgroup contact (`r MlCoeffLatex(lmeMdl = mdlStudent$lmeSlopesAttCoreInt, lmerCI = mdlStudent$lmerSlopesAttCoreIntCI, varName = "OutgroupInteraction")`), this indicates that it is not key need fulfillment in general --- but key need fulfillment during an outgroup contact that predicts more positive outgroup attitudes. This finding is consistent with the results of the previous study, albeit with a slightly weaker effect (likely because of the large number of measurements that did not include an outgroup interaction; For full results see \tblref{tab:robustnessTblLong} and for visual comparison see \fgrref{fig:Robustness}).

@@

In a final step, we again checked whether during the interaction the core situational need remains a meaningful predictor even when taking other fundamental psychological needs into account. We find that the core need fulfillment adds significantly above a model with only the self-determination theory needs (random slopes models; `r studentSdtCompChiSq`). We find that the core need fulfillment explained the most variance in outgroup attitudes after an outgroup contact (`r MlCoeffLatex(lmeMdl = mdlStudentOut$Att$lmeSlopesCoreSdt, lmerCI = mdlStudentOut$Att$lmerSlopesCoreSdtCI, varName = "KeyNeedFullfillment_cwc")`). When compared to the model with only the SDT needs, the core need fulfillment flexibly takes on some of the explained variance of all of the three fundamental needs. However, different from the first study, relatedness (`r MlCoeffLatex(lmeMdl = mdlStudentOut$Att$lmeSlopesCoreSdt, lmerCI = mdlStudentOut$Att$lmerSlopesCoreSdtCI, varName = "Relatedness_cwc")`) and autonomy (`r MlCoeffLatex(lmeMdl = mdlStudentOut$Att$lmeSlopesCoreSdt, lmerCI = mdlStudentOut$Att$lmerSlopesCoreSdtCI, varName = "Autonomy_cwc")`) also predicted positive outgroup attitudes in this larger sample. For full results see \tblref{tab:robustnessTblLong} and \fgrref{fig:Robustness}, and Online Supplementary Material A. This means that also within this second sample, the fulfillment of psychological needs during intergroup contact remained a key predictor of positive outgroup attitudes, even when taking into account several alternative models.

@@

In a final step, we again checked whether during the interaction the core situational need remains a meaningful predictor even when taking other fundamental psychological needs into account. We find that the core need fulfillment adds additional variance above a model with only the self-determination theory needs (random slopes models; `r medicalSdtCompChiSq`). We find that the core need explains the most variance in outgroup attitudes after an outgroup contact (`r MlCoeffLatex(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreSdt, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreSdtCI, varName = "KeyNeedFulfillment_cwc")`). When compared to the model with only the SDT needs, the core need fulfillment flexibly takes on some of the explained variance of all of the three fundamental needs. However, similar to the previous study, in this large sample relatedness (`r MlCoeffLatex(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreSdt, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreSdtCI, varName = "Relatedness_cwc")`), competence (`r MlCoeffLatex(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreSdt, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreSdtCI, varName = "Competence_cwc")`) and autonomy (`r MlCoeffLatex(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreSdt, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreSdtCI, varName = "Autonomy_cwc")`) each also predicted positive outgroup attitudes independently. This being said, the regression coefficient for the core need was three times larger (with all scaling being equal). For full results see \tblref{tab:robustnessTblLong} and \fgrref{fig:Robustness} as well as Online Supplementary Material A.

### 2022-11-07 Robustness-Methods-and-Results.tex
ref fix

To build further confidence in our results, we assessed two additional
models that might offer alternative explanations. First, to ensure that
the effect of core need fulfillment is specific to an actual contact, we
compared the effect to core need fulfillment in situations without an
intergroup contact. For this, we analyzed the generalized situational
core need fulfillment (either during a contact or about the daytime in
general) and tested whether the effect differed during experience
sampling measurements with and without outgroup contacts. We found no
main effect of core need fulfillment (random slopes model; \textit{b} =
-0.10, \textit{t}(1,199) = -2.22, \textit{p} = 0.026,
\textit{95\%CI}{[}-0.18, -0.01{]}) but a significant interaction effect
of core need fulfillment and outgroup contact (\textit{b} = 0.13,
\textit{t}(1,199) = 4.31, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.07, 0.18{]}; also see Table
\ref{tab:robustnessTblLong} and Figure \ref{fig:Robustness}). Together
with a significant main effect of having an outgroup contact, this
indicates that it is not key need fulfillment in general --- but only
key need fulfillment during an outgroup contact that predicts more
positive outgroup attitudes.

In a final step, we controlled for other fundamental psychological needs
during the contact. We focus on the three commonly considered
self-determination needs (SDT): competence, autonomy, and relatedness.
We find that the core need fulfillment adds significantly above a model
with only the self-determination theory needs (random slopes models;
\(\chi^2\)(6, \textit{N} = 21) = 19.60, \textit{p} = 0.003). We also
find that next to relatedness (\textit{b} = 0.10, \textit{t}(361) =
4.04, \textit{p} \textless{} .001, \textit{95\%CI}{[}0.05, 0.14{]}), the
core need explains the most variance in outgroup attitudes after an
outgroup contact (\textit{b} = 0.09, \textit{t}(361) = 2.23, \textit{p}
= 0.026, \textit{95\%CI}{[}0.01, 0.18{]}). When compared to the model
with only the SDT needs, the core need fulfillment flexibly takes on
some of the explained variance of all three fundamental needs
(competence and autonomy needs turning non-significant; all \textit{b}
\textless{} 0.05, all \textit{p} \textgreater{} 0.544). For full results
see see Table \ref{tab:robustnessTblLong}, Figure \ref{fig:Robustness},
and Online Supplementary Material A. There is, thus, considerable
evidence lending confidence to the stability and relevance of
psychological need fulfillment as a predictor of positive outgroup
attitudes for natural intergroup contacts.

@@

Also for Study 2, we checked for alternative models. First, when
considering generalized situational core need fulfillment together with
whether an intergroup contact took place, we find that there is only a
minuscule main effect of core need fulfillment (random slopes model;
\textit{b} = -0.03, \textit{t}(4,849) = -1.34, \textit{p} = 0.181,
\textit{95\%CI}{[}-0.08, 0.02{]}) but a stronger interaction effect of
core need fulfillment and outgroup contact (\textit{b} = 0.06,
\textit{t}(4,849) = 3.03, \textit{p} = 0.002, \textit{95\%CI}{[}0.02,
0.10{]}). Together with a significant main effect of having an outgroup
contact (\textit{b} = 2.88, \textit{t}(4,849) = 3.71, \textit{p}
\textless{} .001, \textit{95\%CI}{[}1.36, 4.40{]}), this indicates that
it is not key need fulfillment in general --- but key need fulfillment
during an outgroup contact that predicts more positive outgroup
attitudes. This finding is consistent with the results of the previous
study, albeit with a slightly weaker effect (likely because of the large
number of measurements that did not include an outgroup interaction; For
full results see Table \ref{tab:intergroupNeedsTblLong} and for visual
comparison see Figure \ref{fig:AllportNeedFulfillment}).

In a final step, we again checked whether during the interaction the
core situational need remains a meaningful predictor even when taking
other fundamental psychological needs into account. We find that the
core need fulfillment adds significantly above a model with only the
self-determination theory needs (random slopes models; \(\chi^2\)(6,
\textit{N} = 108) = 22.90, \textit{p} \textless{} .001). We find that
the core need fulfillment explained the most variance in outgroup
attitudes after an outgroup contact (\textit{b} = 0.08, \textit{t}(823)
= 2.95, \textit{p} = 0.003, \textit{95\%CI}{[}0.03, 0.13{]}). When
compared to the model with only the SDT needs, the core need fulfillment
flexibly takes on some of the explained variance of all of the three
fundamental needs. However, different from the first study, relatedness
(\textit{b} = 0.07, \textit{t}(823) = 4.26, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.04, 0.10{]}) and autonomy (\textit{b} = 0.02,
\textit{t}(823) = 0.93, \textit{p} = 0.353, \textit{95\%CI}{[}-0.02,
0.08{]}) also predicted positive outgroup attitudes in this larger
sample. For full results see Table \ref{tab:intergroupNeedsTblLong},
Figure \ref{fig:Robustness}, and Online Supplementary Material A. This
means that also within this second sample, the fulfillment of
psychological needs during intergroup contact remained a key predictor
of positive outgroup attitudes, even when taking into account several
alternative models.

@@

In a final step, we again checked whether during the interaction the
core situational need remains a meaningful predictor even when taking
other fundamental psychological needs into account. We find that the
core need fulfillment adds additional variance above a model with only
the self-determination theory needs (random slopes models; \(\chi^2\)(6,
\textit{N} = 70) = 100.20, \textit{p} \textless{} .001). We find that
the core need explains the most variance in outgroup attitudes after an
outgroup contact (\textit{b} = 0.15, \textit{t}(1,598) = 3.88,
\textit{p} \textless{} .001, \textit{95\%CI}{[}0.07, 0.22{]}). When
compared to the model with only the SDT needs, the core need fulfillment
flexibly takes on some of the explained variance of all of the three
fundamental needs. However, similar to the previous study, in this large
sample relatedness (\textit{b} = 0.05, \textit{t}(1,598) = 3.31,
\textit{p} \textless{} .001, \textit{95\%CI}{[}0.02, 0.09{]}),
competence (\textit{b} = 0.06, \textit{t}(1,598) = 2.73, \textit{p} =
0.006, \textit{95\%CI}{[}0.02, 0.10{]}) and autonomy (\textit{b} = 0.04,
\textit{t}(1,598) = 2.12, \textit{p} = 0.034, \textit{95\%CI}{[}-0.01,
0.08{]}) each also predicted positive outgroup attitudes independently.
This being said, the regression coefficient for the core need was three
times larger (with all scaling being equal). For full results see Table
\ref{tab:robustnessTblLong} and Figure \ref{fig:Robustness} as well as
Online Supplementary Material A. Across all three studies, psychological
need fulfillment, thus, remained a robust and flexible predictor of
positive outgroup attitudes.

>>>>

To build further confidence in our results, we assessed two additional
models that might offer alternative explanations. First, to ensure that
the effect of core need fulfillment is specific to an actual contact, we
compared the effect to core need fulfillment in situations without an
intergroup contact. For this, we analyzed the generalized situational
core need fulfillment (either during a contact or about the daytime in
general) and tested whether the effect differed during experience
sampling measurements with and without outgroup contacts. We found no
main effect of core need fulfillment (random slopes model; \textit{b} =
-0.10, \textit{t}(1,199) = -2.22, \textit{p} = 0.026,
\textit{95\%CI}{[}-0.18, -0.01{]}) but a significant interaction effect
of core need fulfillment and outgroup contact (\textit{b} = 0.13,
\textit{t}(1,199) = 4.31, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.07, 0.18{]}; also see \tblref{tab:robustnessTblLong}
and \fgrref{fig:Robustness}). Together with a significant main effect of
having an outgroup contact, this indicates that it is not key need
fulfillment in general --- but only key need fulfillment during an
outgroup contact that predicts more positive outgroup attitudes.

In a final step, we controlled for other fundamental psychological needs
during the contact. We focus on the three commonly considered
self-determination needs (SDT): competence, autonomy, and relatedness.
We find that the core need fulfillment adds significantly above a model
with only the self-determination theory needs (random slopes models;
\(\chi^2\)(6, \textit{N} = 21) = 19.60, \textit{p} = 0.003). We also
find that next to relatedness (\textit{b} = 0.10, \textit{t}(361) =
4.04, \textit{p} \textless{} .001, \textit{95\%CI}{[}0.05, 0.14{]}), the
core need explains the most variance in outgroup attitudes after an
outgroup contact (\textit{b} = 0.09, \textit{t}(361) = 2.23, \textit{p}
= 0.026, \textit{95\%CI}{[}0.01, 0.18{]}). When compared to the model
with only the SDT needs, the core need fulfillment flexibly takes on
some of the explained variance of all three fundamental needs
(competence and autonomy needs turning non-significant; all \textit{b}
\textless{} 0.05, all \textit{p} \textgreater{} 0.544). For full results
see see \tblref{tab:robustnessTblLong} and \fgrref{fig:Robustness}, and
Online Supplementary Material A. There is, thus, considerable evidence
lending confidence to the stability and relevance of psychological need
fulfillment as a predictor of positive outgroup attitudes for natural
intergroup contacts.

@@

Also for Study 2, we checked for alternative models. First, when
considering generalized situational core need fulfillment together with
whether an intergroup contact took place, we find that there is only a
minuscule main effect of core need fulfillment (random slopes model;
\textit{b} = -0.03, \textit{t}(4,849) = -1.34, \textit{p} = 0.181,
\textit{95\%CI}{[}-0.08, 0.02{]}) but a stronger interaction effect of
core need fulfillment and outgroup contact (\textit{b} = 0.06,
\textit{t}(4,849) = 3.03, \textit{p} = 0.002, \textit{95\%CI}{[}0.02,
0.10{]}). Together with a significant main effect of having an outgroup
contact (\textit{b} = 2.88, \textit{t}(4,849) = 3.71, \textit{p}
\textless{} .001, \textit{95\%CI}{[}1.36, 4.40{]}), this indicates that
it is not key need fulfillment in general --- but key need fulfillment
during an outgroup contact that predicts more positive outgroup
attitudes. This finding is consistent with the results of the previous
study, albeit with a slightly weaker effect (likely because of the large
number of measurements that did not include an outgroup interaction; For
full results see \tblref{tab:robustnessTblLong} and for visual
comparison see \fgrref{fig:Robustness}).

In a final step, we again checked whether during the interaction the
core situational need remains a meaningful predictor even when taking
other fundamental psychological needs into account. We find that the
core need fulfillment adds significantly above a model with only the
self-determination theory needs (random slopes models; \(\chi^2\)(6,
\textit{N} = 108) = 22.90, \textit{p} \textless{} .001). We find that
the core need fulfillment explained the most variance in outgroup
attitudes after an outgroup contact (\textit{b} = 0.08, \textit{t}(823)
= 2.95, \textit{p} = 0.003, \textit{95\%CI}{[}0.03, 0.13{]}). When
compared to the model with only the SDT needs, the core need fulfillment
flexibly takes on some of the explained variance of all of the three
fundamental needs. However, different from the first study, relatedness
(\textit{b} = 0.07, \textit{t}(823) = 4.26, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.04, 0.10{]}) and autonomy (\textit{b} = 0.02,
\textit{t}(823) = 0.93, \textit{p} = 0.353, \textit{95\%CI}{[}-0.02,
0.08{]}) also predicted positive outgroup attitudes in this larger
sample. For full results see \tblref{tab:robustnessTblLong} and
\fgrref{fig:Robustness}, and Online Supplementary Material A. This means
that also within this second sample, the fulfillment of psychological
needs during intergroup contact remained a key predictor of positive
outgroup attitudes, even when taking into account several alternative
models.

@@

In a final step, we again checked whether during the interaction the
core situational need remains a meaningful predictor even when taking
other fundamental psychological needs into account. We find that the
core need fulfillment adds additional variance above a model with only
the self-determination theory needs (random slopes models; \(\chi^2\)(6,
\textit{N} = 70) = 100.20, \textit{p} \textless{} .001). We find that
the core need explains the most variance in outgroup attitudes after an
outgroup contact (\textit{b} = 0.15, \textit{t}(1,598) = 3.88,
\textit{p} \textless{} .001, \textit{95\%CI}{[}0.07, 0.22{]}). When
compared to the model with only the SDT needs, the core need fulfillment
flexibly takes on some of the explained variance of all of the three
fundamental needs. However, similar to the previous study, in this large
sample relatedness (\textit{b} = 0.05, \textit{t}(1,598) = 3.31,
\textit{p} \textless{} .001, \textit{95\%CI}{[}0.02, 0.09{]}),
competence (\textit{b} = 0.06, \textit{t}(1,598) = 2.73, \textit{p} =
0.006, \textit{95\%CI}{[}0.02, 0.10{]}) and autonomy (\textit{b} = 0.04,
\textit{t}(1,598) = 2.12, \textit{p} = 0.034, \textit{95\%CI}{[}-0.01,
0.08{]}) each also predicted positive outgroup attitudes independently.
This being said, the regression coefficient for the core need was three
times larger (with all scaling being equal). For full results see
\tblref{tab:robustnessTblLong} and \fgrref{fig:Robustness} as well as
Online Supplementary Material A.
